\pkg{casebase}: An Alternative Framework For Survival Analysis and Comparison of Event Rates

### 2021-11-10 analysis/paper/bhatnagar-turgeon-islam-hanley-saarela.Rmd

As discussed in Hanley & Miettinen [-@hanley2009fitting], the key idea behind case-base sampling is to consider the entire study base as an infinite collection of *person-moments*. These person-moments are indexed by both an individual in the study and a time point, and therefore each person-moment has a covariate profile and an outcome status (i.e. whether the event happened) attached to it. By comparing person-moments at which an event occurred with person-moments at which no event occurred, we can extract information about hazards and survival.

@@

>>>>

As discussed in Hanley & Miettinen [-@hanley2009fitting], the key idea behind case-base sampling is to consider the entire study base as an infinite collection of *person-moments*. These person-moments are indexed by both an individual in the study and a time point, and therefore each person-moment has a covariate profile, an exposure status, and an outcome status (i.e. whether the event happened) attached to it. By estimating the probability of an event occurring at a particular person-moment, we can extract information about hazards and survival.

@@

## Variance estimates

The asymptotic normality of $\hat{\theta}$ gives us an efficient way to estimate the variance of various estimators derived from the hazard function. For example, to construct confidence bands around risk functions and survival functions, we can use the following procedure:

  1. Compute $\hat{\theta}$ and its variance-covariance matrix $V(\theta)$.
  2. Sample $B$ times from a multivariate normal $N\left(\hat{\theta}, V(\theta)\right)$.
  3. For each sample, compute the survival function using Equation \ref{eqn:surv} below.
  4. Use the pointwise quantiles of these survival function estimates to construct a pointwise confidence band for the survival function of interest.

This procedure is similar to parametric bootstrap [@efron1994introduction, Section 6.5], but it can be more accurately described as a form of approximate Bayesian computation. As such, the validity of the confidence bands relies on the Bernstein-von Mises theorem [@van2000asymptotic, Chapter 10].

### 2021-11-10 analysis/paper/bhatnagar-turgeon-islam-hanley-saarela.tex

As discussed in Hanley \& Miettinen \citeyearpar{hanley2009fitting}, the
key idea behind case-base sampling is to consider the entire study base
as an infinite collection of \emph{person-moments}. These person-moments
are indexed by both an individual in the study and a time point, and
therefore each person-moment has a covariate profile, an exposure
status, and an outcome status (i.e.~whether the event happened) attached
to it. By comparing person-moments at which an event occurred with
person-moments at which no event occurred, we can extract information
about hazards and survival.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Compute \(\hat{\theta}\) and its variance-covariance matrix
  \(V(\theta)\).
\item
  Sample \(B\) times from a multivariate normal
  \(N\left(\hat{\theta}, V(\theta)\right)\).
\item
  For each sample, compute the survival function using Equation
  \ref{eqn:surv} below.
\item
  Use the pointwise quantiles of these survival function estimates to
  construct a pointwise confidence band for the survival function.
\end{enumerate}

>>>>

As discussed in Hanley \& Miettinen \citeyearpar{hanley2009fitting}, the
key idea behind case-base sampling is to consider the entire study base
as an infinite collection of \emph{person-moments}. These person-moments
are indexed by both an individual in the study and a time point, and
therefore each person-moment has a covariate profile, an exposure
status, and an outcome status (i.e.~whether the event happened) attached
to it. By estimating the probability of an event occurring at a
particular person-moment, we can extract information about hazards and
survival.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Compute \(\hat{\theta}\) and its variance-covariance matrix
  \(V(\theta)\).
\item
  Sample \(B\) times from a multivariate normal
  \(N\left(\hat{\theta}, V(\theta)\right)\).
\item
  For each sample, compute the survival function using Equation
  \ref{eqn:surv} below.
\item
  Use the pointwise quantiles of these survival function estimates to
  construct a pointwise confidence band for the survival function of
  interest.
\end{enumerate}
